\vskip-0.6in Endogenous Politics and the Design of Trade Institutions

### 2016-01-01 draft.tex
remove ex ante section (and proofs); just reference in intro remains

\section{Ex-Ante Lobbying}
\label{sec:exante}

\input{ex_ante}

@@

The discussion in Section~\ref{sec:exante} presumes both that there are no significant interactions between lobbies and that governments are free to choose which sectors are part of a trade agreement, but these decisions are part of a more complicated game that is, in particular, influenced by WTO rules. Taking into account the interactions between multiple lobbies, the influence of WTO rules, and differences in the $\ga$ functions across industries---that is, in how political pressure translates into weight in the policy-making process---seem like important and interesting avenues for further study.

@@

\vskip.2in
\noindent \textbf{\hypertarget{lem1}{Proof of Lemma~\ref{lem:1}}}: \\
In the case of strong bindings, this is immediate from Proposition~\ref{res:strong}. In the case of weak bindings, first examine the first order condition for unilateral tariff setting:
\[
  \frac{\partial W_X(\ga,\tau)}{\partial \tau} = - \frac{5}{49} - \frac{68}{49}\tau +\frac{1}{98}\ga\left[ 16 + 16\tau \right] = 0
\]
Substituting in the inverse of the solution function $\tau^N(\ga) = \frac{8\ga-5}{68-8\ga}$ for $\ga$, this is
\[
  \frac{\partial W_X(\ga,\tau)}{\partial \tau} = - \frac{5}{49} - \frac{68}{49}\tau +\frac{1}{98} \frac{68\tau+5}{8+8\tau} \left[ 16 + 16\tau \right] = 0
\]
This same solution for $\ga$ will be substituted into the first order condition for the joint maximization problem in the ex-post trade agreement game; the only difference is the foreign welfare $W_X^*(\tau) = \frac{25}{98} - \frac{3}{49}\tau + \frac{9}{49}(\tau)^2$ is added to the objective function. The derivative with respect to $\tau$, $- \frac{3}{49} + \frac{18}{49}\tau$ is strictly negative for all non-prohibitive levels of the tariff ($\tau \in \left[0,1/6\right)$) and 0 at the prohibitive tariff. Thus the first order condition for the joint problem is negative at the unilateral tariff; a (weakly) lower tariff must be chosen to achieve the maximal value of this objective function. As demonstrated in Section{sec:objfcn}, the domestic commitment motive may also drive the solution to the free trade corner solution. $\hfill\blacksquare$

@@

In contrast, the last term of the first order condition when bindings are strong is zero since the lobby does not exert any effort at the ex-post stage. The right hand side of Equation~\ref{eq:ea2} then has just the one term, and it can be either greater than or less than 1, so we cannot bound the ex-ante effort level. $\hfill\blacksquare$

>>>>

%\section{Ex-Ante Lobbying}
%\label{sec:exante}

%\input{ex_ante}

@@

%The discussion in Section~\ref{sec:exante} presumes both that there are no significant interactions between lobbies and that governments are free to choose which sectors are part of a trade agreement, but these decisions are part of a more complicated game that is, in particular, influenced by WTO rules.

Taking into account the interactions between multiple lobbies, the influence of WTO rules, and differences in the $\ga$ functions across industries---that is, in how political pressure translates into weight in the policy-making process---seems like an important and interesting avenues for further study.

@@

\begin{comment}
\vskip.2in
\noindent \textbf{\hypertarget{lem1}{Proof of Lemma~\ref{lem:1}}}: \\
In the case of strong bindings, this is immediate from Proposition~\ref{res:strong}. In the case of weak bindings, first examine the first order condition for unilateral tariff setting:
\[
  \frac{\partial W_X(\ga,\tau)}{\partial \tau} = - \frac{5}{49} - \frac{68}{49}\tau +\frac{1}{98}\ga\left[ 16 + 16\tau \right] = 0
\]
Substituting in the inverse of the solution function $\tau^N(\ga) = \frac{8\ga-5}{68-8\ga}$ for $\ga$, this is
\[
  \frac{\partial W_X(\ga,\tau)}{\partial \tau} = - \frac{5}{49} - \frac{68}{49}\tau +\frac{1}{98} \frac{68\tau+5}{8+8\tau} \left[ 16 + 16\tau \right] = 0
\]
This same solution for $\ga$ will be substituted into the first order condition for the joint maximization problem in the ex-post trade agreement game; the only difference is the foreign welfare $W_X^*(\tau) = \frac{25}{98} - \frac{3}{49}\tau + \frac{9}{49}(\tau)^2$ is added to the objective function. The derivative with respect to $\tau$, $- \frac{3}{49} + \frac{18}{49}\tau$ is strictly negative for all non-prohibitive levels of the tariff ($\tau \in \left[0,1/6\right)$) and 0 at the prohibitive tariff. Thus the first order condition for the joint problem is negative at the unilateral tariff; a (weakly) lower tariff must be chosen to achieve the maximal value of this objective function. As demonstrated in Section{sec:objfcn}, the domestic commitment motive may also drive the solution to the free trade corner solution. $\hfill\blacksquare$

@@

In contrast, the last term of the first order condition when bindings are strong is zero since the lobby does not exert any effort at the ex-post stage. The right hand side of Equation~\ref{eq:ea2} then has just the one term, and it can be either greater than or less than 1, so we cannot bound the ex-ante effort level. $\hfill\blacksquare$
\end{comment}
